Cross-Loop Optimization of Arithmetic Intensity for Finite Element Local Assembly

### 2014-10-20 TACO-paper-v2/coffee.tex
Address minor issues in the second round of revisions

We aim to maximize our impact on the platforms that are realistically used
for finite element applications, so we target conventional CPU architectures
rather than GPUs. 
The key limiting factor to the execution on GPUs is the stringent memory requirements. 
Only relatively small problems fit in a GPU memory, and support for
distributed GPU execution in general purpose finite element frameworks is minimal. There has been some research on
adapting local assembly to GPUs (which we mention in Section~\ref{sec:related-work}),
although it differs from ours in several ways, including: (i) not relying on
automated code generation from a domain-specific language (explained next),
(ii) testing only very low order methods, (iii) not optimizing for
cross-loop arithmetic intensity (the goal is rather effective multi-thread
parallelization). In addition, our code transformations would drastically impact the 
GPU parallelization strategy, for example by increasing a thread's working set. 
For all these reasons, a study on extending our work to GPU architectures is 
beyond the scope of this paper. In Section~\ref{sec:generality}, however, we provide 
some intuitions about this research direction.

@@

Experiments were run on a single core of two Intel architectures, a
Sandy Bridge (I7-2600 CPU, running at 3.4GHz, 32KB L1 cache and 256KB
L2 cache) and a Xeon Phi (5110P, running at 1.05Ghz in native mode,
32KB L1 cache and 512KB L2 cache). We have chosen these two
architectures because of the differences in the number of logical
registers and SIMD lanes (16 256-bit registers in the Sandy Bridge,
and 32 512-bit registers in the Xeon Phi), which can impact the
optimization strategy. The \texttt{icc 13.1} compiler was used. On the
Sandy Bridge, the compilation flags used were \texttt{-O2} and
\texttt{-xAVX} for auto-vectorization. On the Xeon Phi, optimization
level \texttt{-O3} was used. Other optimization levels performed, in
general, slightly worse.

@@

COFFEE was developed as a separate, self-contained software module, with
clear input/output interfaces, rather then incorporating it within
PyOP2. This choice was motivated by two critical aspects that
characterize the generality of our research:

>>>>

We aim to maximize our impact on the platforms that are realistically used
for finite element applications, so we target conventional CPU architectures
rather than GPUs. 
The key limiting factor to the execution on GPUs is the stringent memory requirements. 
Only relatively small problems fit in a GPU memory, and support for
distributed GPU execution in general purpose finite element frameworks is minimal. There has been some research on
adapting local assembly to GPUs, which we mention in Section~\ref{sec:related-work},
although it differs from ours in several ways, including: (i) not relying on
automated code generation from a domain-specific language (explained next),
(ii) testing only order 1 methods (while we focus on the range 1-4), (iii) not optimizing for
cross-loop arithmetic intensity (the goal is rather effective multi-thread
parallelization). In addition, our code transformations would drastically impact the 
GPU parallelization strategy, for example by increasing a thread's working set. 
For all these reasons, a study on extending our work to GPU architectures is 
beyond the scope of this paper. In Section~\ref{sec:generality}, however, we provide 
some intuitions about this research direction.

@@

Experiments were run on a single core of two Intel architectures, a
Sandy Bridge (I7-2600 CPU, running at 3.4GHz, 32KB L1 cache and 256KB
L2 cache) and a Xeon Phi (5110P, running at 1.05Ghz in native mode,
32KB L1 cache and 512KB L2 cache). We have chosen these two
architectures because of the differences in the number of logical
registers and SIMD lanes (16 256-bit registers in the Sandy Bridge,
and 32 512-bit registers in the Xeon Phi), which can impact the
optimization strategy. The \texttt{icc 13.1} compiler was used. For what concerns the compilation flags, we selected the best optimization levels on each platform: \texttt{-O2}, and \texttt{-xAVX} for auto-vectorization, were used on the Sandy Bridge; \texttt{-O3} was adopted on the Xeon Phi.

@@

COFFEE was developed as a separate, self-contained software module, with
clear input/output interfaces, rather than incorporating it within
PyOP2. This choice was motivated by two critical aspects that
characterize the generality of our research:
